---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# chariotViz

Visualize the OMOP vocabularies from multiple dimensions using the DiagrammeR package.  

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("meerapatelmd/chariotViz")
```

## Prerequisites  

* OMOP vocabularies loaded into a Postgres instance (see [setupAthena](http://github.com/meerapatelmd/setupAthena)).  
* R.cache R package that caches the query resultsets.    


# How It Works  

## S4 Classes  

```{r,echo=FALSE,message=F,warning=F}
library(DiagrammeR)
DiagrammeR("
  sequenceDiagram;
    Postgres-->>omop.relationships: query; 
    omop.relationships->>nodes.and.edges: divide into;
    nodes.and.edges->>omop.graph: transform to
"
)
```

The raw OMOP data is fetched from the Postgres instance.  

```{r,message=FALSE,warning=FALSE}
library(chariotViz)

omop_rels <- 
  fetch_omop_relationships(version_key = get_version_key())
```

The fetched data belongs to the `omop.relationships` S4 class.  

```{r}
class(omop_rels)
```

The `omop.relationship` class is converted to the `nodes.and.edges` S4 
class. 

```{r}
omop_ne <- 
  create_nodes_and_edges(omop_relationships = omop_rels)

class(omop_ne)
```

From `nodes.and.edges`, an `omop.graph` S4 class object can be constructed.  

```{r}
omop_gr <- 
  construct_graph(omop_ne)

class(omop_gr)
```

The `omop.graph` can then be rendered, but due to its size, `force` must 
be set to TRUE in order to do so.  

```{r}
chariotViz(omop_gr)
```

## Subsetting   

An `omop.relationships` object can be subset using tidyverse-style filtering. 
Here, I am subsetting for all the relationships associated with Cancer Modifier. 
Since the node count is below the threshold set by default in the `chariotViz` function, 
the graph renders automatically and without having to set `force` to TRUE.  

```{r}
cancer_modifier_relationships <- 
  omop_rels %>% 
    filter_omop_relationships(
      vocabulary_id_1 == "Cancer Modifier")

cancer_modifier_ne <- 
  cancer_modifier_relationships %>% 
  create_nodes_and_edges()

cancer_modifier_graph <- 
  cancer_modifier_ne %>% 
  construct_graph()

chariotViz(cancer_modifier_graph)
```


## Adding Pizzaz  

The visualization is nice, but it needs a little oomph. A tooltip and 
visualization attributes can be added to the `nodes.and.edges` class object 
prior to graph construction.  

```{r,echo=FALSE,message=F,warning=F}
DiagrammeR("
  sequenceDiagram;
    Postgres-->>omop.relationships: query; 
    omop.relationships->>nodes.and.edges: divide into;
    nodes.and.edges->>nodes.and.edges: add tooltip;
    nodes.and.edges->>omop.graph: transform to
"
)
```

```{r}
cancer_modifier_relationships <- 
  omop_rels %>% 
    filter_omop_relationships(
      vocabulary_id_1 == "Cancer Modifier")

cancer_modifier_ne <- 
  cancer_modifier_relationships %>% 
  create_nodes_and_edges() %>% 
  add_tooltip()

cancer_modifier_graph <- 
  cancer_modifier_ne %>% 
  construct_graph()

chariotViz(cancer_modifier_graph)
```

In addition to the tooltip, the way the nodes and edges 
appear can also be customized.  

```{r,echo=FALSE,message=F,warning=F}
DiagrammeR("
  sequenceDiagram;
    Postgres-->>omop.relationships: query; 
    omop.relationships->>nodes.and.edges: divide into;
    nodes.and.edges->>nodes.and.edges: add tooltip;
    nodes.and.edges->>nodes.and.edges: add attributes;
    nodes.and.edges->>omop.graph: transform to
"
)
```

```{r}
cancer_modifier_relationships <- 
  omop_rels %>% 
    filter_omop_relationships(
      vocabulary_id_1 == "Cancer Modifier")

cancer_modifier_ne <- 
  cancer_modifier_relationships %>% 
  create_nodes_and_edges() %>% 
  add_tooltip() %>% 
  map_node_attributes() %>% 
  map_edge_attributes()

cancer_modifier_graph <- 
  cancer_modifier_ne %>% 
  construct_graph()

chariotViz(cancer_modifier_graph)
```

## Resizing  

Due to the large number of other relationships for Cancer Modifier, the 
height needs to be increased to make it more legible.  

```{r}
chariotViz(cancer_modifier_graph,
           height = 1500,
           width = 500)
```


## Reconfiguring  

The direction can also be reconfigured.  

```{r}
cancer_modifier_relationships <- 
  omop_rels %>% 
    filter_omop_relationships(
      vocabulary_id_1 == "Cancer Modifier")

cancer_modifier_ne <- 
  cancer_modifier_relationships %>% 
  create_nodes_and_edges() %>% 
  add_tooltip() %>% 
  map_node_attributes() %>% 
  map_edge_attributes()

cancer_modifier_graph <- 
  cancer_modifier_ne %>% 
  construct_graph(attr_theme = "tb")

chariotViz(cancer_modifier_graph,
           width = 2500,
           height = 500)
```

## Adding Example Concepts  

```{r}
set.seed(seed = 1)
cancer_modifier_relationships <- 
  omop_rels %>% 
    filter_omop_relationships(
      vocabulary_id_1 == "Cancer Modifier")

cancer_modifier_ne <- 
  cancer_modifier_relationships %>% 
  create_nodes_and_edges() %>% 
  add_tooltip() %>% 
  map_node_attributes() %>% 
  map_edge_attributes()

cancer_modifier_graph <- 
  cancer_modifier_ne %>% 
  construct_graph() %>% 
  append_concept_examples(sample_size = 3)

chariotViz(cancer_modifier_graph)

chariotViz(cancer_modifier_graph,
           force = TRUE,
           height = 2500,
           width = 700)
```

