---
title: "Getting Started: HemOnc Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started: HemOnc Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error   = FALSE,
  cache.path = file.path(R.cache::getCachePath(),
                         "chariotViz",
                         "vignettes"),
  cache = TRUE
)
```

```{r setup}
library(chariotViz)
```

ChariotViz visualizes all valid Concepts at the Concept Class level by either 
properties represented by the `CONCEPT_RELATIONSHIP` table or by traversing 
direct relationships via the `CONCEPT_ANCESTOR` table.  

OMOP Relationships are fetched from the `CONCEPT_RELATIONSHIP` table with 
corresponding annotations from the `CONCEPT` table.  

```{r,cache=T}
omop_relationships <- 
  fetch_omop(version_key = get_version_key())
```

The `omop.relationships` S4 class object is a dataframe of the query results.  

```{r}
head(omop_relationships@data)
```

This dataframe is then converted to a subset of 2 dataframes of nodes and edges.  

```{r}
omop_nodes_and_edges <- 
  create_nodes_and_edges(omop_relationships = omop_relationships)

head(omop_nodes_and_edges@nodes@data)

head(omop_nodes_and_edges@edges@data)
```

Prior to visualization, a tooltip and attributes are added.  

```{r}
omop_nodes_and_edges <- 
  omop_nodes_and_edges %>% 
  add_tooltip() %>%
  map_node_attributes() %>% 
  map_edge_attributes()
```

The graph is constructed.  

```{r}
omop_graph <- 
  construct_graph(omop_nodes_and_edges)
```

The above is a graph of the entire OMOP Vocabularies in 
its present state.  

```{r}
chariotViz(omop_graph = omop_graph)
```

Due to the large number of nodes, the graph is not rendered, but can 
be forced to do so by setting `force` to TRUE.  

Specifically sets of vocabularies can be isolated by filtering the 
`omop.relationships` class object prior to conversion to nodes and edges.  

Here, we are filtering for relationships within the HemOnc vocabulary only.  

```{r}
hemonc_relationships <- 
  omop_relationships %>% 
  filter_omop_relationships(
    vocabulary_id_1 == "HemOnc",
    vocabulary_id_2 == "HemOnc"
  )
```

The HemOnc graph is constructed after it is filtered for.  

```{r}
hemonc_graph <- 
  create_nodes_and_edges(hemonc_relationships) %>% 
  add_tooltip() %>% 
  map_node_attributes() %>% 
  map_edge_attributes() %>% 
  construct_graph()
```

With a smaller node count, chariotViz will automatically render 
the graph.  

```{r}
chariotViz(hemonc_graph,
           width = 2000,
           height = 2000)
```


The fillcolor of the nodes can be remapped.  

```{r}
hemonc_graph2 <- 
  hemonc_graph %>% 
  remap_fillcolor_by_concept_class()

chariotViz(hemonc_graph2,
           width = 2000,
           height = 2000)
```

# Relationships  

The relationships can also be filtered based on 3 
classes of relationships: Taxonomy, Mapping, or Lateral. These classes 
are not defined by OHDSI and are specific to the `chariot` and `chariotViz` 
packages.  

## Taxonomy  

Taxonomy relationships are relationships that represent a hierarchy (`Is a` and `Subsumes`).  

```{r}
hemonc_taxonomy_relationships <- 
  hemonc_relationships %>% 
  filter_for_taxonomy_relationships() 

hemonc_taxonomy_graph <- 
  create_nodes_and_edges(hemonc_taxonomy_relationships) %>% 
  add_tooltip() %>% 
  map_node_attributes() %>% 
  map_edge_attributes() %>% 
  construct_graph() %>% 
  remap_fillcolor_by_concept_class()

chariotViz(hemonc_taxonomy_graph,
           height = 500,
           width = 500)
```

## Mapping  

Mapping relationships are relationships that represent a mapping from a source to 
target, as in the case of an ETLing into the OMOP CDM. Since we filtered for 
concepts within HemOnc, the filter returns zero relationships since the basis 
of these relationships is mapping between different vocabularies.    

```{r}
hemonc_mapping_relationships <- 
  hemonc_relationships %>% 
  filter_for_mapping_relationships() 

hemonc_mapping_relationships
```

To view HemOnc's mapping relationships, we can re-filter the original object for 
HemOnc on side 1 only.  

```{r}
all_hemonc_relationships <- 
  omop_relationships %>% 
  filter_omop_relationships(
    vocabulary_id_1 == "HemOnc")

hemonc_mapping_relationships <- 
  all_hemonc_relationships %>% 
  filter_for_mapping_relationships() 

hemonc_mapping_relationships
```

We can now see how HemOnc maps to other OMOP vocabularies.  

```{r}
hemonc_mapping_graph <- 
  create_nodes_and_edges(hemonc_mapping_relationships) %>% 
  add_tooltip() %>% 
  map_node_attributes() %>% 
  map_edge_attributes() %>% 
  construct_graph() 

chariotViz(hemonc_mapping_graph,
           height = 500,
           width = 500)
```

## Lateral  

Lateral relationships are relationships that are all other relationships. 
Returning to the relationships within HemOnc, the lateral relationships 
are filtered for.  

```{r}
hemonc_lateral_relationships <- 
  hemonc_relationships %>% 
  filter_for_lateral_relationships()

hemonc_lateral_graph <- 
  create_nodes_and_edges(hemonc_lateral_relationships) %>% 
  add_tooltip() %>% 
  map_node_attributes() %>% 
  map_edge_attributes() %>% 
  construct_graph() 

chariotViz(hemonc_lateral_graph,
           height = 2000,
           width = 2000)
```

With the relationships within and outside of HemOnc, the `all_hemonc_relationships` 
returns the following graph.  

```{r}
hemonc_lateral_relationships <- 
  all_hemonc_relationships %>% 
  filter_for_lateral_relationships()

hemonc_lateral_graph <- 
  create_nodes_and_edges(hemonc_lateral_relationships) %>% 
  add_tooltip() %>% 
  map_node_attributes() %>% 
  map_edge_attributes() %>% 
  construct_graph() 

chariotViz(hemonc_lateral_graph,
           height = 2000,
           width = 2000)
```

## Foreign  

To retrieve any relationships outside of HemOnc instead, the original `omop_relationships` 
object must be filtered once more, but this time with exclusion of HemOnc in 
the filter.  

```{r}
foreign_hemonc_relationships <- 
  omop_relationships %>% 
  filter_omop_relationships(
    vocabulary_id_1 == "HemOnc",
    vocabulary_id_2 != "HemOnc") %>% 
  filter_for_lateral_relationships() 

foreign_hemonc_lateral_graph <- 
  create_nodes_and_edges(foreign_hemonc_relationships) %>% 
  add_tooltip() %>% 
  map_node_attributes() %>% 
  map_edge_attributes() %>% 
  construct_graph()

chariotViz(foreign_hemonc_lateral_graph)
```

Filtering for Taxonomy relationships allows visualization of how ancestry of 
a vocabulary relates to other vocabularies. Here, it is evident that only 
HemOnc's `Component Class` is a parent to RxNorm and RxNorm Extension Concept Classes.  

```{r}
foreign_hemonc_relationships <- 
  omop_relationships %>% 
  filter_omop_relationships(
    vocabulary_id_1 == "HemOnc",
    vocabulary_id_2 != "HemOnc") %>% 
  filter_for_taxonomy_relationships() 

foreign_hemonc_taxonomy_graph <- 
  create_nodes_and_edges(foreign_hemonc_relationships) %>% 
  add_tooltip() %>% 
  map_node_attributes() %>% 
  map_edge_attributes() %>% 
  construct_graph()

chariotViz(foreign_hemonc_taxonomy_graph)
```

To view all possible foreign relationships, the filter can be lifted.    

```{r}
foreign_hemonc_relationships <- 
  omop_relationships %>% 
  filter_omop_relationships(
    vocabulary_id_1 == "HemOnc",
    vocabulary_id_2 != "HemOnc") #%>% 
#  filter_for_lateral_relationships() 

foreign_hemonc_graph <- 
  create_nodes_and_edges(foreign_hemonc_relationships) %>% 
  add_tooltip() %>% 
  map_node_attributes() %>% 
  map_edge_attributes() %>% 
  construct_graph()

chariotViz(foreign_hemonc_graph)
```

# Visualizing Concepts  

Samples of concepts along with a tooltip containing the concept's attribute 
values can be appended to the graph with the `append_concept_examples` function.  

```{r}
foreign_hemonc_graph %>% 
  append_concept_examples(sample_size = 3) %>% 
  chariotViz(force = TRUE,
             width = 1500)
```

The count of the samples can be adjusted.  

```{r}
foreign_hemonc_graph %>% 
  append_concept_examples(sample_size = 5) %>% 
  chariotViz(force = TRUE,
             width = 1500)
```

```{r}
foreign_hemonc_graph %>% 
  append_concept_examples(sample_size = 10) %>% 
  chariotViz(force = TRUE,
             height = 2000)
```

